---
title: "Lab 04 - Data Visualization"
author: "Qiushi"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---
```{r}
library(lubridate)
library(tidyverse)
library(data.table)
```


## Step 1. Read in the data
First download and then read in with data.table:fread()
```{r read-data, cache = TRUE}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

## Step 2. Prepare the data

Remove temperatures less than -17C. Make sure there are no missing data in the key variables coded as 9999, 999, etc.
```{r}
met <- met[temp > -17][elev==9999.0, elsv := NA]
```

Generate a date variable using the functions as.Date()
```{r create-ymd, cache = TRUE}
met <- met[ , ymd := as.Date(paste(year, month, day, sep = "-"))]
```

Using the data.table::week function, keep the observations of the first week of the month.
```{r}
met[, table(week(ymd))]
met <- met[week(ymd) == 31]
```

Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
```{r check-max}
met[, .(
  temp = max(temp, na.rm = T),
  rh = max(rh, na.rm = T),
  wind.sp = max(wind.sp, na.rm = T),
  vis.dist = max(wind.sp, na.rm = T),
  dew.point = max(dew.point, na.rm = T),
  lat = max(lat, na.rm = T),
  lon = max(lon, na.rm = T),
  elev = max(elev, na.rm = T)
)]
```
Great! No more 9999s in our dataset.

```{r check-avg}
met_avg <- met[, .(
  temp = mean(temp, na.rm = T),
  rh = mean(rh, na.rm = T),
  wind.sp = mean(wind.sp, na.rm = T),
  vis.dist = mean(wind.sp, na.rm = T),
  dew.point = mean(dew.point, na.rm = T),
  lat = mean(lat, na.rm = T),
  lon = mean(lon, na.rm = T),
  elev = mean(elev, na.rm = T)
), by = "USAFID"]
```

Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
```{r}
met_avg[, region := fifelse(lon >= -98 & lat > 39.71, "NE",
                fifelse(lon < -98 & lat > 39.71, "NW",
                fifelse(lon < -98 & lat <= 39.71, "SW", "SE")))
]
table(met$region)
```

Create a categorical variable for elevation as in the lecture slides
```{r}
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]
```

## Step 3. Make Violin plots of dew point temp by region
```{r}
met_avg[!is.na(region)] %>%
  ggplot() +
  geom_violin(mapping = aes(x = 1, y= dew.point, color = region, fill = region)) +
  facet_wrap(~ region, nrow = 1)
```

The highest dew point temperatures are reported in the southeast.

```{r}
met_avg[!is.na(region) & !is.na(wind.sp)] %>%
  ggplot() +
  geom_violin(mapping = aes(x = 1, y = wind.sp, color = region, fill = region)) +
  facet_wrap(~ region, nrow = 1)
```

The highest wind speeds are reported in the northeast.

## Step 04. Use geom_point with geom_smooth to examine the association between dew point temperature and wind speed by region
Colour points by region
Make sure to deal with NA category
Fit a linear regression line by region
Describe what you observe in the graph

```{r scatterplot-dewpoint-wind.sp}
met_avg[!is.na(region) & !is.na(wind.sp)] %>%
  ggplot(mapping = aes(x = wind.sp, y = dew.point, color = region)) +
  geom_point(mapping = aes(color = region)) +
  geom_smooth(method = lm, mapping = aes(linetype = region)) +
  facet_wrap(~ region, nrow = 2)
```
The dew point is positively associated with the wind speed in the southeast.

## Step 5. Use geom_bar to create barplots of the weather stations by elevation category coloured by region
```{r}

```


## Step 6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars
```{r}
met_avg[!is.na(dew.point)] %>%
  ggplot(mapping = aes(x = region, y = dew.point)) + 
    stat_summary(fun.data = mean_sdl, geom = "errorbar")
```


## Step 7. Make a map showing the spatial trend in relative h in the US
```{r}
# Generating a color palette
library(leaflet)
rh.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)
rh.pal
```

```{r}
top10rh <- met_avg[ rank(-rh) <= 10]
```


```{r}
rhmap <- leaflet(met_avg) %>% 
  # The looks of the Map
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~lat, lng=~lon,
                                                  # HERE IS OUR PAL!
    label = ~paste0(rh), color = ~ rh.pal(rh),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=rh.pal, values=met_avg$rh,
          title='Relative Humid', opacity=1)
rhmap
```


