---
title: "Assignment 01"
author: "Qiushi"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

Import packages
```{r}
library(data.table)
library(leaflet)
library(dplyr)
```

## Step 1.read in data and chek for issues

Reading in the 2004 data:
```{r read_data_1}
pm_04 <- data.table::fread("ad_viz_plotval_data_2004.csv")
pm_04
```
Reading in the 2019 data:
```{r read_data_2}
pm_19 <- data.table::fread("ad_viz_plotval_data_2019.csv")
pm_19
```
Check dimention:
```{r}
dim(pm_04)
dim(pm_19)
```

Check headers:
```{r}
head(pm_04)
head(pm_19)
```

Check footers:
```{r}
tail(pm_04)
tail(pm_19)
```

Check variable names and types:
```{r}
str(pm_04)
str(pm_19)
```
Check issues:
```{r}
summary(pm_04$`Daily Mean PM2.5 Concentration`)
summary(pm_19$`Daily Mean PM2.5 Concentration`)
table(is.na(pm_04))
table(is.na(pm_19))
```


Summary: The maximum PM2.5 concentration in 2004 is much more larger than that in 2019. There are many missing values.

## Step 2. Combine the two years of data into one data frame. Use the Date variable to create a new column for year, which will serve as an identifier. Change the names of the key variables so that they are easier to refer to in your code.

Combine the two years of data into one data frame:
```{r combine_data}
pm <- rbind(pm_04, pm_19)
```



Use the Date variable to create a new column for year, which will serve as an identifier. Change the names of the key variables so that they are easier to refer to in your code.
```{r new_column}
pm <- mutate(pm, year = factor(rep(c(2004, 2019), c(nrow(pm_04), nrow(pm_19))))) %>%
  rename(PM = `Daily Mean PM2.5 Concentration`)
```


## Step 3. Create a basic map in leaflet() that shows the locations of the sites (make sure to use different colors for each year). Summarize the spatial distribution of the monitoring sites.
```{r leaflet}
library(leaflet)
year.pal <- colorFactor(c("blue","red"), domain = pm$year)

leaflet(pm) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(lat = ~SITE_LATITUDE, lng = ~SITE_LONGITUDE,
                   color = ~year.pal(year),
                   fillOpacity = 1, radius = 2) %>% 
  addLegend("bottomleft", pal = year.pal, values = pm$year, title = "year")

```



Summary: Monitoring sites are densely distributed in coastal areas, and sparsely distributed in the inland areas. The number of monitoring sites in 2019 is larger than that in 2004.

## Step 4. Check for any missing or implausible values of PM2.5 in the combined dataset. Explore the proportions of each and provide a summary of any temporal patterns you see in these observations.

Check missing or implausible values.
```{r}
summary(pm$PM)
mean(pm$PM < 0)
mean(is.na(pm$PM))
```
There is no missing value of PM2.5. But there are 0.39% implausible values.

Remove lines with implausible values (PM2.5 < 0).
```{r replace_implausible values}
pm <- pm[(PM > 0) | (PM = 0)]
```


## Step 5. Explore the main question of interest at three different spatial levels. Create exploratory plots (e.g. boxplots, histograms, line plots) and summary statistics that best suit each level of data. Be sure to write up explanations of what you observe in these data.

### State view (entire California)
```{r}
library(ggplot2)
ggplot(data = pm, mapping = aes(year, log2(PM))) + geom_boxplot()
```

```{r}
with(pm, tapply(PM, year, summary))
```


Summary: PM2.5 of 2004 is much more larger than that of 2019 in California. In other words, the air pollution of 2019 is less than that of 2004. Also the variation in 2019 is smaller than that in 2004.

### County View
```{r}
pm_county <- pm[, list(PM, COUNTY, year)]
```

```{r}
pm_county <- pm_county[ , .(mean(PM, na.rm = T)), by = .(COUNTY, year)]
pm_county
```
Line plot may be a good choice.
```{r}
qplot(Year, V1, data = mutate(pm_county, Year = as.numeric(as.character(year))),
      color = factor(COUNTY), 
      geom = c("point", "line")) +
  ylab('PM2.5')
  
```

Summary: Most counties in California decreased PM2.5 from 2004 to 2016. Only a few of them increased the value. Riversde had the most significant decrease in PM2.5, which was from 18.424356 to 8.506640 ug/m3.

### Site View in Los Angeles

```{r}
pm_LA <- 
  pm %>%
  filter(COUNTY == 'Los Angeles') %>%
  collect()

pm_LA <- 
  pm_LA %>%
  select(PM, 'Site Name', year)
pm_LA <- pm_LA[ , .(mean(PM, na.rm = T)), by = .(`Site Name`, year)]
pm_LA
```

Remove line with blank "Site Name"
```{r}
pm_LA <- pm_LA %>% filter(`Site Name` != "") %>% collect()
pm_LA
```
Remove lines with site that only appear once.
```{r}
sites = intersect(x = pm_LA[year == 2004]$`Site Name`, y = pm_LA[year == 2019]$`Site Name`)
sites

pm_LA <- 
  pm_LA %>%
  filter(pm_LA$`Site Name` %in% sites) %>% 
  collect()
pm_LA
```



Try bar chart.
```{r}
ggplot(data = pm_LA, mapping = aes(x = V1, y = `Site Name`, fill = year)) + 
         geom_bar(stat = "identity", position = "dodge") +
  xlab("PM2.5") + ylab("")
```

Summary: There was decrease in PM2.5 in all sites in Los Angeles, based on the bar chart. Many sites, for example Azusa, had decreased about 50%. Lebec had the best air quality in both 2004 and 2019. The air quality of Los Angeles-North Main Street was worst.


