---
title: "Lab 03"
author: "Qiushi"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1. Read in the data
First download and then read in with data.table:fread()

```{r}
if (!file.exists("met_all.gz")) {
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
}

met <- data.table::fread("met_all.gz")
```

## Step 2. Check the dimensions, headers, footers. How many columns, rows are there?

```{r}
dim(met)
```
```{r}
head(met)
```
```{r}
tail(met)
```

## Step 3. Take a look at the variables.
```{r}
str(met)
```

## Step 4. Take a closer look at the key variables.
```{r}
table(met$year)
```

```{r}
table(met$day)
```

```{r}
table(met$hour)
```

```{r}
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
```

Replace elevation with 9999 as NA.
```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

The weather station with the highest elevation is at `r max(met$elev, na.rm=TRUE)` meter.

Remove the temps below -40 degree Celsius.
```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

## 5. Check the data against an external data source.
```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

## Step 6. Calculate summary statistics

```{r}
met[elev == max(elev, na.rm = TRUE), summary(wind.sp)]
met[elev == max(elev, na.rm = TRUE), summary(temp)]
```


```{r}
met[elev == max(elev, na.rm = TRUE), .(
  temp_wind = cor(temp, wind.sp, use="complete"),
  temp_hour = cor(temp, hour, use="complete"),
  wind.sp_day = cor(wind.sp, day, use="complete"),
  wind.sp_hour = cor(wind.sp, hour, use="complete")
)]
```

## Step 7. Exploratory graphs
```{r}
hist(met$elev, breaks=100)
```
```{r}
hist(met$temp)
```
```{r}
hist(met$wind.sp)
```
```{r}
if (!require(leaflet)) {
  install.packages("leaflet")
  library(leaflet)
}
elev <- met[elev == max(elev)]
```

```{r}
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

```{r}
if (!require(lubridate)) {
  install.packages("leaflet")
  library(lubridate)
}
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```

```{r}
elev <- elev[order(date)]
head(elev)
```
```{r}
plot(elev$date, elev$temp, type='l')
```
```{r}
plot(elev$date, elev$wind.sp, type='l')
```

