---
title: "Assignment 02 - Data Viz and Wrangling"
author: "Qiushi"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

Import libraries.
```{r}
library(data.table)
library(dplyr)
library(ggplot2)
```


### Data Wrangling

Read in data.
```{r read-data}
individual <- read.csv("chs_individual.csv")
regional <- read.csv("chs_regional.csv")
```

Merge data.
```{r}
total_data <- 
  merge(x = individual,
        y = regional,
        by = "townname",
        all.x = TRUE,
        all.y = FALSE)
total_data
```

## Step 1

Identify and remove duplicates.
```{r}
total_data <- unique(total_data)
dim(total_data)
```
Rows matched.

Impute data using the average within the variables “male” and “hispanic”.
```{r impute-data}
total_data <- data.table(total_data)
total_data[ , bmi := fcoalesce(bmi, mean(bmi, na.rm = TRUE)), by = .(male, hispanic)]
```

## Step 2

Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24).
```{r categorize}
total_data[ , obesity_level := fifelse(bmi < 14, "underweight", 
                                       fifelse(bmi <= 22, "normal", 
                                               fifelse(bmi <= 24, "overweight", "obese")))]
total_data
```

To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.
```{r min-max-total}
summary_table <- 
  total_data[ , .(
  Minimum = min(bmi, na.rm = TRUE),
  Maximum = max(bmi, na.rm = TRUE),
  Total_Number = length(bmi)
), by = obesity_level]

knitr::kable(summary_table)

```

## Step 3

Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.”
```{r smoke_gas_exposure}
total_data[ , smoke_gas_exposure := fifelse((smoke == 0) & (gasstove == 0), "no",
  fifelse((smoke == 1) & (gasstove == 0), "smoke",
    fifelse((smoke == 0) & (gasstove == 1), "gasstove", "both")))]
total_data
```

## Step 4

Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”

Table grouped by town.
```{r table_town}
table_town <-
  total_data[ , .(
  FEV_avg = mean(fev, na.rm = TRUE),
  FEV_sd = sd(fev, na.rm = TRUE),
  Asthma_indicator = sum(asthma, na.rm = TRUE) / sum(!is.na(asthma))
), by = townname]
knitr::kable(table_town)
```

Table grouped by sex
```{r table_sex}
table_sex <-
  total_data[ , .(
  FEV_avg = mean(fev, na.rm = TRUE),
  FEV_sd = sd(fev, na.rm = TRUE),
  Asthma_indicator = sum(asthma, na.rm = TRUE) / sum(!is.na(asthma))
), by = male]
knitr::kable(table_sex)
```

Table grouped by obesity level.
```{r table_obesity}
table_obesity <-
  total_data[ , .(
  sample_size = 
  FEV_avg = mean(fev, na.rm = TRUE),
  FEV_sd = sd(fev, na.rm = TRUE),
  Asthma_indicator = sum(asthma, na.rm = TRUE) / sum(!is.na(asthma))
), by = obesity_level]
knitr::kable(table_obesity)
```

Table grouped by smoke_gas_exposure.
```{r table_smoke_gas_exposure}
table_smoke_gas_exposure <-
  total_data[ , .(
  FEV_avg = mean(fev, na.rm = TRUE),
  FEV_sd = sd(fev, na.rm = TRUE),
  Asthma_indicator = sum(asthma, na.rm = TRUE) / sum(!is.na(asthma))
), by = smoke_gas_exposure]
knitr::kable(table_smoke_gas_exposure)
```

### Looking at the Data (EDA)

## 1. Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.

```{r BMIvsFEV}
total_data[!is.na(bmi) & !is.na(fev)] %>%
  ggplot(mapping = aes(x = bmi, y = fev)) + 
  geom_point(mapping = aes(color = townname)) + 
  geom_smooth(method = lm) +
  facet_wrap(~ townname, nrow = 4)
```

## 2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.

Stacked histogram by BMI category.
```{r stacked-histogram-1}
total_data[!is.na(obesity_level) & !is.na(fev)] %>%
  ggplot(aes(fev, fill = obesity_level)) +
  geom_histogram(binwidth = 20)
```

Stacked histogram by smoke/gas exposure
```{r stacked-histogram-2}
total_data[!is.na(smoke_gas_exposure) & !is.na(fev)] %>%
  ggplot(aes(fev, fill = smoke_gas_exposure)) +
  geom_histogram(binwidth = 20)
```

## 3. Barchart of BMI by smoke/gas exposure.

```{r barchart}
bmi_exposure <- total_data[, list(bmi, smoke_gas_exposure)]

group_by(bmi_exposure[!is.na(smoke_gas_exposure) & !is.na(bmi)], smoke_gas_exposure) %>%
  summarize_each(funs = mean) %>%
  ggplot() + 
  geom_col(mapping = aes(y = bmi, x = smoke_gas_exposure, colour = smoke_gas_exposure, fill = smoke_gas_exposure))
```

## 4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.

Statistical summary graphs of FEV by BMI category.
```{r}
total_data[!is.na(fev) & !is.na(obesity_level)] %>%
  ggplot(mapping = aes(x = obesity_level, y = fev)) + 
    stat_summary(fun.data = mean_sdl, geom = "pointrange") +
    stat_summary(fun.data = mean_sdl, geom = "errorbar") 
```

Statistical summary graphs of FEV by smoke/gas exposure category.
```{r}
total_data[!is.na(fev) & !is.na(smoke_gas_exposure)] %>%
  ggplot(mapping = aes(x = smoke_gas_exposure, y = fev)) + 
    stat_summary(fun.data = mean_sdl, geom = "pointrange") +
    stat_summary(fun.data = mean_sdl, geom = "errorbar") 
```

## 5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.

```{r leaflet}
library(leaflet)
pm.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain = total_data$pm25_mass)

leaflet(total_data) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addCircles(
    lat = ~lat, lng=~lon,
    label = ~paste0(pm25_mass), color = ~ pm.pal(pm25_mass),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  addLegend('bottomleft', pal = pm.pal, values = total_data$pm25_mass,
          title = 'PM2.5 mass', opacity=1)
```

## 6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.

```{r}
total_data[!is.na(pm25_mass) & !is.na(fev)] %>%
  ggplot(mapping = aes(x = pm25_mass, y = fev)) + 
  geom_point() +
  geom_smooth(method = lm)
```

